---
title: "Data Wrangling With R - Week 4"
author: "Nitin"
date: "November 8, 2016"
output: 
    html_document:
       theme: lumen 
---

#Synopsis
This file encompasses the Week 4 Homework of the Data Wrangling with R class. This week concentrates on exploring the gapminder dataset availble on CRAN

```{r message=FALSE}
#Load relevant libraries
library(gapminder)
library(dplyr)
library(ggplot2)
library(ggthemes)
```


```{r include=FALSE}
?gapminder
```

#Source Code  
country: factor with 142 levels  
continent: factor with 5 levels  
year: ranges from 1952 to 2007 in increments of 5 years  
lifeExp: life expectancy at birth, in years  
pop: population  
gdpPercap: GDP per capita  

#Data Description

```{r include=FALSE}
str(gapminder_unfiltered)
```

###Basic Data Structure and Summary Statistics
The data set includes 3313 observations of 6 variables  
The variables include country, continent, year, lifeExp, pop and gdpPercap  

```{r }
summary(gapminder_unfiltered)
#psych::describe(gapminder_unfiltered)
#Hmisch::describe(gapminder_unfiltered)
```

##Exploratory Data Analysis {.tabset}

###What is the distribution of GDP per capita across all countries?

```{r}
gapminder_unfiltered %>% 
  filter(year == 2007) %>%
  ggplot(aes(gdpPercap))+geom_histogram(binwidth   = 2000, color = "white") +
  labs(title = "Histogram of GDP Per Capita") +
  labs(x="GDP Per Capita" , y = "Frequency") +
  geom_vline(aes(xintercept=mean(gdpPercap)),
            color="grey", linetype="dashed", size=0.5) 
```

We observe that the GDP Per Capita is heavily right skewed.  

###How do the distributions differ across the different continents?

```{r}
gapminder_unfiltered %>% 
  filter(year == 2007) %>%
  ggplot(aes(x = continent, y = gdpPercap, fill = continent))+geom_boxplot() 
```

We observe that the distributions are quite different across continents with Asia showcasing the maximum range and Africa with the least range.  
Europe seems to have the highes mean while Asia showcases the largest GDP Per capita  

###What are the top 10 countries with the largest GDP per capita?

```{r}
gapminder_unfiltered %>% 
  filter(year == 2007) %>%
  arrange(desc(gdpPercap)) %>%
  top_n(n = 10, gdpPercap) %>%
  ggplot(aes(x = reorder(country, gdpPercap), y = gdpPercap, fill = continent)) + geom_bar(stat = "identity", width = 0.7) + coord_flip() +
labs(title = "GDP Per Capita Over the Ages (India)") +
  labs(x = "Country", y = "GDP Per Capita") 
```

We observe that primarily Asian countries represent the top 10 countries with highest GDP Per capita


###Plot the GDP per capita for your country of origin for all years available

```{r}
gapminder_unfiltered %>% 
  filter(country == "India") %>%
  ggplot(aes(x = year, y = gdpPercap)) + 
  geom_line() +
  labs(title = "GDP Per Capita Over the Ages (India)") +
  labs(x = "Year", y = "GDP Per Capita") +
  geom_point()
```

We observe an exponential growth over the past few decades

###What was the percent growth (or decline) in GDP per capita in 2007?

```{r}
growth_rate <- gapminder_unfiltered %>% 
  filter(country == "India") %>% mutate(Change = (gdpPercap - lag(gdpPercap))/(lag(gdpPercap)) * 100) %>%
  select(Country = country,Year = year, Change)
growth_rate
```

We observe a growth of 40.38%  

###What has been the historical growth (or decline) in GDP per capita for your country?

```{r}
mean(growth_rate$Change, na.rm = TRUE)
```

Historically India has been growing at 15%

